---
title: "Comparing Mean Length for Female Lizards of Two Different Species"
author: "Joe Walderman"
date: "12/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(tidyverse)
library(here)
library(janitor)
library(broom)
library(stargazer)
library(GGally)
library(kableExtra)
library(effsize)
```

## Intro and Data

In this analysis I use [data](https://doi.org/10.6073/pasta/ff37b4cace16a9943575f3fd7067064e) collected by David Lightfoot et al in the northern Chihuahuan Desert. The data contains observations on species, sex, recapture status, total length in mm, weight in grams, and whether tail is broken or whole among other information. For this analysis I created a data visualization and table to compare total length between two species and performed a T-test and Cohen's d test to compare means and measure effect size. Further, I use histograms and QQ-plots to check on the assumptions of the hypothesis testing. These tests look to answer the question is there significant difference in total length difference in mean total length between female lizards for the two species.

```{r}
lizards <- read.csv(here("data", "lizards.csv")) %>% 
  clean_names()
```

```{r}
lizards_subset <- lizards %>% 
  select(spp, sex, total_length)
```

## Analysis

```{r}
lizards_f <- lizards_subset %>% 
  filter(sex == "F",
         spp %in% c("CNUN", "CNTE"))

ggplot(data = lizards_f, aes(x = spp, y = total_length)) +
  geom_boxplot(aes(fill = spp)) +
  stat_summary(
    fun = mean,
    geom = "point"
    ) +
  labs(x = "Species",
       y = "Length (mm)",
       title = "Lizard Length by Species") +
  scale_x_discrete(labels= c("CNTE"= "Colorado Checkered Lizard", "CNUN"="Desert Grassland Lizard")) +
  theme(legend.position = "none") 
```

**Figure 2.** Total lentgh (mm) observations for female lizards of species Colorado Checkered and Desert Grassland. The width of the plots indicate the number of observations for a particular weight. Box endpoints indicate the 25th and 75th percentile values; the black line and black point within the box indicate the median and mean value for each sex, respectively. Data: Lightfoot, 2017.




**Table 1.** Descriptive statistics (mean, standard deviation, and sample size) for Colorado Checkered Lizards and Desert Grassland Lizards. Data: Lightfoot, 2017.
```{r}
lizards_table <- lizards_f %>% 
  group_by(spp) %>% 
  summarize(mean_weight = round(mean(total_length, na.rm = TRUE),2),
            sd_weight = round(sd(total_length, na.rm = TRUE),2),
            sample_size = n()) %>%
  mutate(spp = case_when(
    spp == "CNTE" ~ "Colorado Chekered Lizards",
    spp == "CNUN" ~ "Desert Grassland Lizard"
  )) %>% 
  kable(col.names = c("Species", "Mean Total Lenth (mm)", "Standard Deviation (mm)", "Sample Size"), digits = 2) %>% 
  kable_styling(bootstrap_options = "striped", full_width = FALSE)
  
lizards_table
```



```{r}
#Pulling vectors of weight for male and female juvenile hares
lizards_CNTE <- lizards_f %>% 
  filter(spp == "CNTE") %>% 
  pull(total_length)

lizards_CNUN<- lizards_f %>% 
  filter(spp == "CNUN") %>% 
  pull(total_length)

# hist(lizards_CNTE) # seems skewed to the right
# qqnorm(lizards_CNTE) #not very linear
# hist(lizards_CNUN) #also seems kind of skewed right
# qqnorm(lizards_CNUN) # more linear

# Getting means and sds for lizard total length by species
CNTE_mean <- mean(lizards_CNTE, na.rm = TRUE)
CNUN_mean <- mean(lizards_CNUN, na.rm = TRUE)
CNTE_sd <- sd(lizards_CNTE, na.rm = TRUE)
CNUN_sd <- sd(lizards_CNUN, na.rm = TRUE)
```

```{r}
# T-test for ...
lizards_ttest <- t.test(lizards_CNTE, lizards_CNUN, na.rm = TRUE)
lizards_ttest_tidy <- tidy(lizards_ttest)

# Cohen's d effect size for...
lizards_effsize <- cohen.d(lizards_CNTE, lizards_CNUN, na.rm = TRUE)
```

## Summary

The mean total length of the observed female Colorado Checkered lizards (`r round(CNTE_mean,1)` mm) was greater than the mean total length of female Desert Grassland Lizards (`r round(CNUN_mean,1)` mm).The Welch's two sample t-test yields a p-value of t(`r round(lizards_ttest_tidy$parameter,2)`) = `r round(lizards_ttest_tidy$statistic,2)`, p < 0.001) which indicates there is a negligible chance of randomly selecting two samples from populations with the same mean that are this different by chance. They are almost certainly from populations with different mean lizard lengths. The absolute difference in the means (`r round(CNTE_mean,2) - round(CNUN_mean,2)` mm), and the large effect size (Cohen's *d* = `r round(lizards_effsize$estimate,2)`), along with the p-value contribute to the conclusion that there is significant difference in total length difference in mean total length between female lizards for the two species.


## Data Citation

Lightfoot, D. 2017. Lizard pitfall trap data (LTER-II, LTER-III) ver 36. Environmental Data Initiative. https://doi.org/10.6073/pasta/ff37b4cace16a9943575f3fd7067064e (Accessed 2020-12-07).

